% !Mode:: "TeX:UTF-8"

%%% 说明: 此部分需要自己填写的内容:  (1) 中文摘要及关键词 (2) 英文摘要及关键词
%--------------------------------------------------------------------------
%%%页眉设置
%
%\pagestyle{fancyplain}
%\fancyhf{}  %清除以前对页眉页脚的设置
%\renewcommand{\headrulewidth}{0.5pt}
%\fancyfoot[C]{-\,\thepage\,-}
%\fancypagestyle{plain}
%{
%\fancyhead[CE]{武汉大学博士学位论文}
%\fancyhead[CO]{}
%\renewcommand{\headrulewidth}{0.5pt}
%\fancyfoot{}                                    % clear all footer fields
%\fancyfoot[C]{-\,\thepage\,-}
%}

%%%%%%%%%%%%%%%%%%%%%%%
%%% ------------ 中文摘要 ---------------%%%
%%%%%%%%%%%%%%%%%%%%%%%
\begin{cnabstract}
    深度学习的快速发展为人工智能领域带来了突破性进展, 然而神经网络内部的特征提取与作用机制缺乏合理的可解释性, 制约着该技术的进一步发展与可信应用. 特征学习理论 (Feature Learning Theory, FLT) 作为深度学习领域的新兴基础理论, 致力于揭示神经网络从数据中提取并利用特征的内在规律, 为深度学习的可解释性与可信性提供理论依据. 
    
    本文围绕特征学习理论的核心科学问题, 系统研究了神经网络在不同任务场景下的特征学习机制. 首先, 在通用任务场景下, 本文聚焦对抗训练 (Adversarial Training) 中的特征塌缩现象, 通过大量实证研究发现了单纯形压缩现象：在对抗训练的最终阶段, 特征会收敛到单纯形等角紧框架 (Simplex ETF) 结构, 但其几何尺寸较标准训练显著缩小, 且压缩程度与扰动半径呈负相关关系. 本文从理论层面阐释了对抗训练引发该现象的作用机制, 证明了单纯形范数与扰动半径之间的负相关关系, 为对抗训练下的特征演化提供了理论解释. 
    
    其次, 在图像任务场景下, 本文针对基于重构的异常检测 (Reconstruction-Based Anomaly Detection, RBAD) 问题, 建立了分析卷积自编码器特征学习的理论框架. 通过引入锥集直觉, 本文刻画了神经网络如何从正常训练数据中提取特征, 以及学习到的特征在重构阶段的作用机制. 理论分析表明, 在正常数据上训练的两层卷积自编码器以高概率无法准确重构异常, 从而为RBAD的有效性提供了理论支撑. 实验验证了理论分析的准确性, 并在真实世界数据集上展示了清晰的特征学习过程. 
    
    本文的研究成果不仅丰富了特征学习理论的内涵, 也为对抗训练和异常检测等实际应用场景提供了理论指导, 对推动深度学习技术的可解释性和可信应用具有重要意义. 
\end{cnabstract}
\vspace{1em}\par\vfill

%%%--------- 关键词 -------- %%%
\cnkeywords{神经网络, 特征学习, 对抗训练, 异常检测}

%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%
%%% ------------ 英文摘要 ---------------%%%
%%%%%%%%%%%%%%%%%%%%%%%

\begin{enabstract}
    The rapid development of deep learning has brought breakthrough progress to the field of artificial intelligence. However, the lack of reasonable interpretability regarding the feature extraction and mechanism within neural networks constrains the further development and trustworthy application of this technology. Feature Learning Theory (FLT), as an emerging fundamental theory in the field of deep learning, is dedicated to revealing the intrinsic laws of how neural networks extract and utilize features from data, providing theoretical foundations for the interpretability and trustworthiness of deep learning.
    
    This thesis systematically investigates the feature learning mechanisms of neural networks in different task scenarios, focusing on the core scientific problems of feature learning theory. First, in the general task scenario, this thesis focuses on the neural collapse phenomenon in adversarial training (AT). Through extensive empirical research, we discover the simplex compression phenomenon: in the terminal phase of adversarial training, features converge to a simplex equiangular tight frame (Simplex ETF) structure, but its geometric size is significantly smaller than that of standard training, and the compression degree exhibits a negative correlation with the perturbation radius. This thesis theoretically explains the mechanism by which adversarial training induces this phenomenon, proves the negative correlation between the simplex norm and the perturbation radius, and provides a theoretical explanation for feature evolution under adversarial training.
    
    Second, in the image task scenario, this thesis addresses the reconstruction-based anomaly detection (RBAD) problem and establishes a theoretical framework for analyzing feature learning in convolutional autoencoders. By introducing the cone set intuition, this thesis characterizes how neural networks extract features from normal training data and the mechanism of learned features in the reconstruction phase. Theoretical analysis shows that a two-layer convolutional autoencoder trained on normal data cannot accurately reconstruct anomalies with high probability, thereby providing theoretical support for the effectiveness of RBAD. Experiments validate the accuracy of the theoretical analysis and demonstrate a clear feature learning process on real-world datasets.
    
    The research results of this thesis not only enrich the connotation of feature learning theory but also provide theoretical guidance for practical application scenarios such as adversarial training and anomaly detection, which is of great significance for promoting the interpretability and trustworthy application of deep learning technology.
\end{enabstract}
\vspace{1em}\par\vfill

%%%------ 英文关键词 ------- %%%
\enkeywords{Neural Networks, Feature Learning, Adversarial Training, Anomaly Detection}


