\bibitem{CCB2022Benign} CAO Y, CHEN Z X, BELKIN M, et al. Benign Overfitting in Two-layer Convolutional Neural Networks[C]//Proceedings of the 36th Annual Conference on Neural Information Processing Systems. Red Hook: Curran Associates, Inc., 2022.

\bibitem{AL2021Feature} ALLEN-ZHU Z, LI Y Z. Feature Purification: How Adversarial Training Performs Robust Deep Learning[C]//Proceedings of the 62nd Annual Symposium on Foundations of Computer Science. Los Alamitos: IEEE Computer Society, 2021.

\bibitem{CCL2025Prevalence} CAO Y, CHEN Y B, LIU W W. Prevalence of Simplex Compression in Adversarial Deep Neural Networks[J]. Proceedings of the National Academy of Sciences of the United States of America, 2025.

\bibitem{AL2023Towards} ALLEN-ZHU Z, LI Y Z. Towards Understanding Ensemble, Knowledge Distillation and Self-Distillation in Deep Learning[C]//Proceedings of the 11th International Conference on Learning Representations. OpenReview.net, 2023.

\bibitem{WL2021Toward} WEN Z X, LI Y Z. Toward understanding the feature learning process of self-supervised contrastive learning[C]//Proceedings of the 34th International Conference on Machine Learning. PMLR, 2021: 11014-11024.

\bibitem{HCW2024Graph} HUANG W, CAO Y, WANG H N, et al. Graph Neural Networks Provably Benefit from Structural Information: A Feature Learning Perspective[C]//Proceedings of the 12th International Conference on Learning Representations. OpenReview.net, 2024.

\bibitem{LWY2024Benign} LU M, WU B N, YANG X D, et al. Benign Oscillation of Stochastic Gradient Descent with Large Learning Rates[C]//Proceedings of the 12th International Conference on Learning Representations. OpenReview.net, 2024.

\bibitem{ZCL2023Benefits} ZOU D F, CAO Y, LI Y Z, et al. The Benefits of Mixup for Feature Learning [C]//Proceedings of the 37th International Conference on Machine Learning. PMLR, 2023: 27265-27280.

\bibitem{CHZ2023Towards} CHEN Y Q, HUANG W, ZHOU K W, et al. Towards Understanding Feature Learning in Out-of-Distribution Generalization[C]//Proceedings of the 37th International Conference on Machine Learning. PMLR, 2023: 4827-4837.

\bibitem{CDL2024Understanding} CHEN Z X, DENG Y H, LI Y Z, et al. Understanding Transferable Representation Learning and Zero-shot Transfer in CLIP[C]//Proceedings of the 12th International Conference on Learning Representations. OpenReview.net, 2024.

\bibitem{CL2026Two-Layer} CHEN Y B, LIU W W. Two-Layer Convolutional Autoencoders Trained on Normal Data Provably Detect Unseen Anomalies[C]//Proceedings of the 14th International Conference on Learning Representations. OpenReview.net, 2026.

\bibitem{RHW1986Learning} RUMELHART D, HINTON G E, WILLIAMS R J. Learning Internal Representations by Error Propagation[M]//Parallel Distributed Processing: Explorations in the Microstructure of Cognition, Vol. 1: Foundations. Cambridge: MIT Press, 1986: 318-362. ISBN 026268053X.

\bibitem{BCV2013Representation} BENGIO Y, COURVILLE A, VINCENT P. Representation Learning: A Review and New Perspectives[C]//Proceedings of the 2nd International Conference on Learning Representations. OpenReview.net, 2013.

\bibitem{CB2020Implicit} CHIZAT L, BACH F. Implicit Bias of Gradient Descent for Wide Two-layer Neural Networks Trained with the Logistic Loss[C]//Proceedings of the 33rd Conference on Learning Theory. PMLR, 2020: 1273-1316.

\bibitem{PHD2020prevalence} PAPYAN V, HAN X, DONOHO D. Prevalence of neural collapse during the terminal phase of deep learning training[J]//Proceedings of the National Academy of Sciences of the United States of America. 2020.

\bibitem{MMSTV2017towards} MA̧DRY A, MAKELOV A, SCHMIDT L, TSIPRAS D, VLADU A. Towards Deep Learning Models Resistant to Adversarial Attacks[C]//International Conference on Learning Representations. 2017.

\bibitem{arXiv:math/0301135} STROHMER T, HEATH R. Grassmannian Frames with Applications to Coding and Communication[R]//CoRR.  arXiv:math/0301135.

\bibitem{arXiv:1504.00253} FICKUS M, MIXON D G. Tables of the existence of equiangular tight frames[R]//CoRR. arXiv:1504.00253

\bibitem{ZYJXGJ2019theoretically} ZHANG H, YU Y, JIAO J, XING E, GHAOUI L, JORDAN M I. Theoretically Principled Trade-off between Robustness and Accuracy[C]//International Conference on Machine Learning. 2019.
